|

### 2023-06-15 paper/article.qmd
fixed cross-references and minor writing issues

To test these hypotheses, I combine two datasets.
The first was created through content analysis and contains measures of framing in reports about a large number of events in UK mainstream newspapers from 1992 to the end of 2017.
The second one contains information about protest events.
These data are matched and the hypotheses are tested through multilevel logistic regression models, with the presence or absense of a frame as dependant, and the above described factors that are assumed to shape protest reporting as independent variables.

@@

Table @tab:outlets presents the selection of newspaper outlets, strategically chosen to contain variety on the two important factors: 
ideology and type of newspaper in the UK
Additionally the selection was made to include the outlets with the highest circulation figures.
I retrieved articles from LexisNexis, utilizing keywords such as "protest" and "demonstration" and several several misspellings.
A pilot ensured the comprehensiveness of these keywords by testing a range of synonyms like "march", "rally", or "direct action", which, however, did not yield any additional relevant articles.
The articles underwent a processing phase using the R-package LexisNexisTools [@gruberlnt2021] which consisted of removing duplicate entries, items discussing non-domestic protests, and false positives.
To maintain theoretical integrity, we opted to limit our data to domestic protests, as studies have indicated notable differences in the coverage of non-domestic protests [@boyleAdherenceProtestParadigm2012].
This led to the final dataset, comprising 27,496 newspaper articles.

```{r outlets}
#| label: outlets
outlets <- tibble::tribble(
  ~Type,        ~Ideology,  ~Outlet,                 ~"Available from", ~"Circulation (2005 -- 2017)",
  "Tabloid",    "Left",    "The Mirror",             1995,              "1,720,000 -- 692,295",
  "Tabloid",    "Right",   "The Sun",                2000,              "3,273,000 -- 1,602,320",
  "Tabloid",    "Right",   "Daily Mail",             1992,              "2,426,000 -- 1,442,924",
  "Broadsheet", "Left",    "The Guardian",           1992,              "366,645 -- 153,431",
  "Broadsheet", "Left",    "The Independent",        1992,              "263,595 -- NA",
  "Broadsheet", "Right",   "Financial Times",        1992,              "419,386 -- 190,046",
  "Broadsheet", "Right",   "Daily Telegraph",        1992,              "460,585 -- 907,329",
  "Broadsheet", "Right",   "The Times",              1992,              "679,190 -- 440,736"
)

footnotes = c(
  "Source for circulation numbers: http://www.magforum.com/papers/nationals.htm",
  "No figures available for The Independent in 2017 as the outlet moved to online-only publication in 2016."
)

outlets |>  
  knitr::kable(format = "latex",
        booktabs = TRUE,
        linesep = "",
        caption = "Selection of Newspaper Outlets for Data Collection",
        longtable = FALSE,
        escape = FALSE) |>  
  kableExtra::add_footnote(footnotes, notation = "alphabet", threeparttable = TRUE)
```

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from @gruberTrouble2023.
This results in one binary variable per frame: which takes the value 1 if a frame is present and 0 otherwise.
One article can contain several frames, or none.
In short, I frames were identified by coding the frame elements found in @entman1993's definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected artiocle and combining the coded elements to frames using factor analysis.
They were then coded using supervised machine learning models trained on the manually coded material.

@@

This article tests the relation between framing of protest and eight independent variable identified from the literature above:
*Protest Goal*, *Protester Violence*, *State Repression*, *Newspaper Ideology*, *Newspaper Type*, *Ideological Divide*, *Days Since Start*, and *Year of Protest*.
*Newspaper Ideology* were obtained through the meta information of the coded news articles *Newspaper Type*.
For all other variables the *Mass Mobilization Project*'s [@ClarkMMPdata2019] dataset on protest events was employed.
This dataset was chosen for pratical reasons as it contains the variables neccesary and covers the chosen time frame (1992-2017).
@ClarkMMPdata2019 use the keywords "Protest", "Demonstration", "Riot", and "Mass Mobilization" on the *LexisNexis* newspaper archive, selecting "All News" as the source.
For the selected time frame in the UK, the *Mass Mobilization Project* lists 471 events between 1992 and the end of 2017, as shown in @fig-MMPyear.

```{r}
#| label: fig-MMPyear
#| fig-cap: "Number of Protest Events in the UK per Year per Clark and Regan (2019)"
readRDS("../misc/protest_event_data_uk.Rds")  |> 
  mutate(year = lubridate::floor_date(date_start, "year"))  |> 
  group_by(year)  |> 
  count()  |> 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = jbg_pal(sel = "grey")) +
  scale_x_date(labels = lubridate::year, breaks = "year",
               expand = c(0, 0)) +
  #scale_y_continuous(labels = scales::percent, limits = c(0, 0.7)) +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```

To merge the two datasets, I employed a two-staged strategy:
first, I selected all articles in the newspaper dataset that were published no earlier than seven days before an event started and no later than 30 days after it ended;
then I used the variables "Protest Locations", "Protest Group Identity" and "Protester Demands" to extract suitable keywords [see Appendix @app-eventdata].
I then used keywords to filter the articles selected in the first stage:
articles were discarded if they did not feature at least one of the keywords.
This selection was manually validated and overlaps between events were separated by hand.

For the variable *ideological divide*, additional manual coding was necessary:
An ideological divide is present when the newspaper ideology is right-leaning and the protest ideology is coded as left --- or the other way around.
The newspaper ideology is known; however, I had to code the ideology of the protest myself based on the information in the MMP dataset and in other articles about an event.^[See Appendix @app-eventdata for the specific coding.]
Note, that since "right"- and "left"-leaning are concepts which are often disputed, I tried to assess whether a protest aligns closer with the main UK right-wing or left-wing party (i.e., closer to Labour or Conservative politics), as this is how British newspapers are usually divided [@kuhnPoliticsMediaBritain2007].

Using the variables listed in @tab-variables, I fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectivly consitute panel dataset since observations must be seen as nested and not independent.
It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets.
However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016].
Therefore, I used newspaper-year as the Level 2 variable ($n = 180$).
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->

```{r variables}
#| label: variables
variables <- rio::import("../misc/variables.csv")  |> 
  mutate(across(everything(), ~gsub('"', '', .x, fixed = TRUE)))  |> 
  select(-`Levels-long`, -Mean, -Category)

@@

\newpage
# Appendix {.unnumbered}

>>>>

To test these hypotheses, I combine two datasets.
The first was created through content analysis and contains measures of framing in reports about a large number of events in UK mainstream newspapers from 1992 to the end of 2017.
The second one contains information about protest events.
These data are matched and the hypotheses are tested through multilevel logistic regression models.
@tbl-variables shows the dependant and independant variables of these models --- one for the presence or absense of each frame.
How this data was sourced is explained below.

```{r variables}
#| label: tbl-variables
#| tbl-cap: Model Variables
variables <- rio::import("../misc/variables.csv")  |> 
  mutate(across(everything(), ~gsub('"', '', .x, fixed = TRUE)))  |> 
  select(-`Levels-long`, -Mean, -Category)

@@

@tbl-outlets presents the selection of newspaper outlets, strategically chosen to contain variance on the two important factors: 
ideology and type of newspaper in the UK
Additionally the selection was made to include the outlets with the highest circulation figures.
I retrieved articles from LexisNexis, utilizing the keywords "protest" and "demonstration" and several several misspellings.
A pilot ensured the comprehensiveness of these keywords by testing a range of synonyms like "march", "rally", or "direct action", which, however, did not yield any additional relevant articles.
The articles underwent a processing phase using the R-package LexisNexisTools [@gruberlnt2021] which consisted of removing duplicate entries, items discussing non-domestic protests, and false positives.
To maintain theoretical integrity, I opted to limit our data to domestic protests, as studies have indicated notable differences in the coverage of non-domestic protests [@boyleAdherenceProtestParadigm2012].
This led to the final dataset, comprising 27,496 newspaper articles.

```{r outlets}
#| label: tbl-outlets
#| tbl-cap: Selection of Newspaper Outlets for Data Collection
outlets <- tibble::tribble(
  ~Type,        ~Ideology,  ~Outlet,                 ~"Available from", ~"Circulation (2005 -- 2017)",
  "Tabloid",    "Left",    "The Mirror",             1995,              "1,720,000 -- 692,295",
  "Tabloid",    "Right",   "The Sun",                2000,              "3,273,000 -- 1,602,320",
  "Tabloid",    "Right",   "Daily Mail",             1992,              "2,426,000 -- 1,442,924",
  "Broadsheet", "Left",    "The Guardian",           1992,              "366,645 -- 153,431",
  "Broadsheet", "Left",    "The Independent",        1992,              "263,595 -- NA",
  "Broadsheet", "Right",   "Financial Times",        1992,              "419,386 -- 190,046",
  "Broadsheet", "Right",   "Daily Telegraph",        1992,              "460,585 -- 907,329",
  "Broadsheet", "Right",   "The Times",              1992,              "679,190 -- 440,736"
)

footnotes = c(
  "Source for circulation numbers: http://www.magforum.com/papers/nationals.htm",
  "No figures available for The Independent in 2017 as the outlet moved to online-only publication in 2016."
)
library(gt)
outlets |>
  gt() |>
  tab_footnote(
    footnote = "Source for circulation numbers: http://www.magforum.com/papers/nationals.htm",
    locations = cells_column_labels(columns = "Circulation (2005 -- 2017)")
  ) |>
  tab_footnote(
    footnote = "No figures available for The Independent in 2017 as the outlet moved to online-only publication in 2016.",
    locations = cells_body(columns = "Circulation (2005 -- 2017)", rows = 5L)
  )
```

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from @gruberTrouble2023.
In short, frames were identified by coding the frame elements found in @entman1993's definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected articles and combining the coded elements to frames using factor analysis.
This results in one binary variable per frame, which takes the value 1 if a frame is present and 0 otherwise.
Each article can contain several frames, or none.
After training supervised machine learning models trained on the manually coded random sample, all remaining articles were coded using these models.

@@

This article tests the relation between framing of protest and eight independent variable identified from the literature above:
*Protest Goal*, *Protester Violence*, *State Repression*, *Newspaper Ideology*, *Newspaper Type*, *Ideological Divide*, *Days Since Start*, and *Year of Protest*.
*Newspaper Ideology* were obtained through the meta information of the coded news articles *Newspaper Type*.
For all other variables the *Mass Mobilization Project*'s [@ClarkMMPdata2019] dataset on protest events was employed.
This dataset was chosen for pratical reasons as it contains the variables neccesary and covers the chosen time frame (1992-2017).
@ClarkMMPdata2019 use the keywords "Protest", "Demonstration", "Riot", and "Mass Mobilization" on the *LexisNexis* newspaper archive, selecting "All News" as the source.
For the selected time frame in the UK, the *Mass Mobilization Project* lists 471 events between 1992 and the end of 2017, as shown in @fig-mmpyear.

```{r}
#| label: fig-mmpyear
#| fig-cap: "Number of Protest Events in the UK per Year per Clark and Regan (2019)"
readRDS("../misc/protest_event_data_uk.Rds")  |> 
  mutate(year = lubridate::floor_date(date_start, "year"))  |> 
  group_by(year)  |> 
  count()  |> 
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = jbg_pal(sel = "grey")) +
  scale_x_date(labels = lubridate::year, breaks = "year",
               expand = c(0, 0)) +
  #scale_y_continuous(labels = scales::percent, limits = c(0, 0.7)) +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```

To merge the two datasets, I employed a two-staged strategy:
first, I selected all articles in the newspaper dataset that were published no earlier than seven days before an event started and no later than 30 days after it ended;
then I used the variables "Protest Locations", "Protest Group Identity" and "Protester Demands" to extract suitable keywords [see Appendix].
I then used keywords to filter the articles selected in the first stage:
articles were discarded if they did not feature at least one of the keywords.
This selection was manually validated and overlaps between events were separated by hand.

For the variable *ideological divide*, additional manual coding was necessary:
An ideological divide is present when the newspaper ideology is right-leaning and the protest ideology is coded as left --- or the other way around.
The newspaper ideology is known; however, I had to code the ideology of the protest myself based on the information in the MMP dataset and in other articles about an event.^[See Appendix for the specific coding.]
Note, that since "right"- and "left"-leaning are concepts which are often disputed, I tried to assess whether a protest aligns closer with the main UK right-wing or left-wing party (i.e., closer to Labour or Conservative politics), as this is how British newspapers are usually divided [@kuhnPoliticsMediaBritain2007].

Using the variables listed in @tbl-variables, I fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectivly consitute panel dataset since observations must be seen as nested and not independent.
It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets.
However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016].
Therefore, I used newspaper-year as the Level 2 variable ($n = 180$).
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->

@@

\newpage
# Appendix {#sec-eventdata .unnumbered}
